Dotted Suffix Trees\\A Structure for Approximate Text Indexing

### 2005-07-29 paper/construction.tex
Polish it a bit Still not proved O(nh^k)!

Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}.

@@

To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs, filtering the leafs labeled one) and make this the root's error tree. Then, for every other node, remove the current error tree (it has a level too few). Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label).

>>>>

Note that these conditions are very similiar to the conditions for the root and both can be implemented as shown in Figure~\ref{algo:copy-subtree}. This is an expression of Lemma~\ref{lemma:suffix-error}.

@@

To construct the (k+1)-error tree from the k-error tree, make a copy of the tree as above (adjusting leafs and filtering the leafs labeled one) and make this the new root error tree. Then, for every other node, remove the current error tree (it has a level too few). Finally, for every node except the root, construct its error tree as above (adjusting leafs, filtering the leafs labeled one or which do not correspond to positions in the suffix link's label). These error trees will now have the correct number of nodes.

### 2005-07-29 paper/space.tex
Polish it a bit Still not proved O(nh^k)!

\[ \hat{h} = \frac{1}{n} \sum_{x \mathit{node of \mathcal{T}}} \mathit{Height}(x), n \mathrm{being the number of nodes} \]
\end{definition}

@@

\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\frac{1}{n}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = n\hat{h} \qed \]

We will reason by induction, starting with the case of a one-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree.

@@

We define $\mathit{Weight}(\overline{w})$ as above. The total number of nodes for all the error trees in the dotted tree is therefore equal to $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w})$ where $\mathcal{T}$ is the original suffix tree. Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.

The difference between a k-error and a (k+1)-error tree is that all the

>>>>

\[ \hat{h} = \frac{1}{|\mathcal{T}|} \sum_{x \in \mathit{\mathcal{T}}} \mathit{Height}(x) \]
\end{definition}

@@

\[ \mathit{Weight}(\mathcal{T}) = \sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\frac{1}{|\mathcal{T}|}\sum_{x \in \mathcal{T}}\mathit{Height}(x) = |\mathcal{T}|\hat{h} \qed \]

We will reason by induction, starting with the case of a zero-error dotted tree to show, that the bound on the number of nodes in an k-error dotted tree can also be expressed as $O(n\hat{h}^k)$ where $\hat{h}$ is the average height of the suffix tree. The base case is a suffix tree and is obvious.

@@

%Assume the k-error suffix tree has $O(h\hat{h}^k)$ nodes.

%We define $\mathit{Weight}(\overline{w})$ as above. For any node, the number of leafs in its error tree will be $O(\mathit{Weight}(\overline{w}))$. The total number of nodes for all the error trees in the dotted tree is therefore equal to $O(\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}))$ where $\mathcal{T}$ is the original suffix tree. Given the lemma above, $\sum_{\overline{w} \in \mathcal{T}}\mathit{Weight}(\overline{w}) = O(n\hat{h})$.

%The difference between a k-error and a (k+1)-error tree is that all the

### 2005-07-29 paper/structure.tex
Polish it a bit Still not proved O(nh^k)!

Each node in a $\Sigma^+$ tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
\end{definition}

@@

For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$. In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously. Note that for each node, the tree below it is the $\Sigma^+$ tree of the suffixes starting at the positions in its position set.

In a leaf, the position set is a singleton, and we label the leaf by its element.

@@

For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a $\Sigma^+$ tree.

@@

The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}.

@@

\begin{definition}[k-error dotted tree]
We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree.
\end{definition}

>>>>

Each node in a Patricia tree has a path leading to it which forms a string. If the node $n$ has the leading path $w$, we shall also refer to $n$ as $\overline{w}$.
\end{definition}

@@

For example, in the node $\overline{issi}$ in the suffix tree of Figure~\ref{fig:mississipi-0}, the occurence set is $\{5, 2\}$ and its position set is $\{6, 9\}$. In a sense, one can say that being at node $\overline{issi}$ is being at positions 6 and~9 simultaneously. Note that for each node, the tree below it is the Patricia tree of the suffixes starting at the positions in its position set.

In a leaf, the occurence set is a singleton, and we label the leaf by its element.

\begin{lemma}[Position set at the suffix node]\label{lemma:suffix-error}
Given two nodes $\overline{aw}$ and $\overline{w}$ (connected by a suffix link, labeled $a$), if one takes position set of $\overline{w}$, substracts one to each element, one obtains a superset of the position set of $\overline{w}$. The items shared by both sets are those positions of the string which contain an $a$.
\end{lemma}

The lemma is fairly obvious given that the position set of $\overline{aw}$ contains all the positions where $aw$ occurs which will be exactly one minus those positions where $w$ occurs following $a$.

@@

For example, in the above mentioned node $\overline{issi}$, the error tree is formed by taking the strings starting at positions $\{7, 10\}$ (ie \putstring{sipi\$} and \putstring{i\$}) in a Patricia tree.

@@

\begin{figure}
\includegraphics{figures/mississipi-1}
\caption{1-error dotted tree for \putstring{mississipi}}%
\label{fig:mississipi-1}
\end{figure}

The one-error dotted tree for \putstring{mississipi} is shown in Figure~\ref{fig:mississipi-1}. The nodes are connected to their error trees by diagonal links.

@@

\begin{definition}[k-error dotted tree]
We define a k-error dotted tree as the tree obtained by adding error trees to each node in the (k-1)-error dotted tree which does not already contain one.
\end{definition}
